---
title: "01-bf_data_basics"
format: pdf
---

From PHMRC to BFL: the Absolute Basics

This vignette explains, at the most basic level, what information BFL actually needs, starting from a raw PHMRC-style dataset.

This is written to teach both the user and the author what is going on.

# 1 . What a PHMRC dataset really is

A PHMRC-style dataset is just a table.

-   Each **row** represents **one individual**.

-   Each **column** represents **some attribute** of that individual.

At minimum, the table contains **three types of columns**:

## (1) A site/region identifier

Example:

```         
site = "AP", "Mexico", "Bohol"
```

This column tells us **which individuals belong together**.

Without a site column, BFL does not make sense.

## (2) A cause-of-death label

```         
cause = "Stroke", "Malaria", "TB"
```

This is a **categorical variable** describing the outcome we want to model.

Eventually, models convert these labels into integers:

```         
Stroke  → 1
Malaria → 2
TB      → 3
```

But conceptually, this column answers:

> What happened to this person?

## (3) Symptom indicators

Example columns:

```         
fever, cough, chest_pain, headache, ...
```

Each row has values like:

```         
0 / 1
TRUE / FALSE
Yes / No
```

These columns describe observable evidence.

# 2. First transformation: split by site

BFL never works on the full table at once.

The first operation is always:

> Split the dataset by site.

For a site "AP":

```{r}
phmrc_AP <- phmrc[phmrc$site == "AP", ]
```

Now phmrc_AP is still just a table, but it contains only one region.

# 3. Second transformation: extract X and Y

From the site-specific table, we extract exactly two objects.

### **(A) Feature matrix X**

We take only the symptom columns and convert them to a numeric matrix:

```{r}
X_AP <- as.matrix(phmrc_AP[, symptom_columns])
```

This matrix has shape:

```         
X_AP : N x P
```

-   N = number of individuals at the site

-   P = number of symptoms

This is the only individual-level data used later for prediction.

### **(B) Cause vector Y**

We take the cause column and encode it as intergers:

```{r}
Y_AP <- as.integer(factor(phmrc_AP$cause))
```

This vector has length N and values in:

```{1,2, ..., C_AP}
```

where C_AP is the nubmer of causes observed at that site.

# 4. At this point, PHMRC is gone

After extraction, the original table is no longer needed.

Everything downstream is built from:

```{r}
X_AP # N x P matrix
Y_AP # length-N intger vector
```

No person IDs.

No site column.

No raw table.

# 5. What LCVA does to (X,Y)

LCVA does not care about individual rows.

It learns relationships between symptoms and causes.

From (X_AP, Y_AP), LCVA produces a matrix:

```         
posterior_phi_AP: P x C_AP
```

where:

```         
posterior_phi_AP[p,c] = P(symptom p|cause c)
```

interpretation:

-   Rows = symptoms

-   Columns = causes

-   Values = probabiltiies

This matrix summarizes all individuals at the site.

# 6. What survives and what is discarded

After LCVA:\
\
**Discarded**

-   Individual rows

-   Person idenities

-   Raw symptom patterns

-   Raw cause labels

\
**Kept**

-   posterior_phi_AP (P x C Matrix)

-   The list of cause IDs used

-   The ordering of symptoms

This is why LCVA is often described as:

> Turning people into beliefs.

# 7. What BFL actually needs

For each site, BFL needs only:

```{r}
list(
  posterior_phi = posterior_phi_AP,
  cause_ids = c(1,2,...,C_AP)
)
```

That is the entire input to the global BFL model.

No individual-level data enters BFL.

# 8. What happens with multiple sites

Suppose we have three sites:

```         
AP      → posterior_phi_AP (P × C_AP)
Mexico  → posterior_phi_MX (P × C_MX)
Bohol   → posterior_phi_BO (P × C_BO)
```

These matrices may:

-   have different number of caues

-   share some causes

-   miss others entirely

BFL's job is to:

1.  Align causes across sites
2.  Combine these matrices
3.  Infer global parameters that summarize all sites

BFL never sees individual rows.

# 9. One sentence summary

> **PHMRC → (X, Y) → LCVA → posterior_phi → BFL**

Everythign else in the package exists to support this transformation.

# 10. Why this matters

Understanding this flow makes it clear:

-   why privacy is preserved

-   why LCVA and BFL are seperate

-   why BFL aggregates matrices, not people

-   why prediction happen locally

Once this mental model is clear, the API design becomes obvious.
